\documentclass{article}

\usepackage{amsmath, amssymb}

\begin{document}
\section*{Properties of Greedy Algorithm}
Problems that can be solved by greedy algorithms have two main properties:
\begin{description}
    \item[Optimal Substructure] The optimal solution to the problem contains
        within it optimal solutions to subproblems.
    \item[Greedy Choice Property] A global optimal solution can be arrived at by
        selecting a local optimal solution.
\end{description}

Difference from DP: No overalapping subproblems.

\section*{Activity Selection Problem}

\begin{description}
    \item[Input] Set \(S\) of \(n\) activities \(a_1, a_2, \ldots, a_n\).
        \begin{itemize}
            \item \(s_i\) = start time of activity \(i\)
            \item \(f_i\) = finish time of activity \(i\)
        \end{itemize}
    \item[Output] Subset \(A\) of maximum number of compatible activities.
\end{description}
Two activities are compatible if they do not overlap.

\begin{tabular}{c|ccccccccc}
    i       & 1 & 2 & 3 & 4 & 5 & 6  & 7  & 8  & 9  \\
    \hline
    \(s_i\) & 1 & 2 & 4 & 1 & 5 & 8  & 9  & 11 & 13 \\
    \(f_i\) & 3 & 5 & 7 & 8 & 9 & 10 & 11 & 14 & 16 \\
\end{tabular}

Assume activites are sorted by finishing times.
\begin{equation*}
    f_1 \leq f_2 \leq \ldots \leq f_n
\end{equation*}

Suppose an optimal solution includes activity \(a_k\).

\begin{itemize}
    \item This generates two subproblems
          \begin{itemize}
              \item Selecting from \(a_1, \ldots, a_{k-1}\) activites compatible
                    with one another, and that finish before \(a_k\) starts
                    (compatible with \(a_k\))
              \item Selecting from \(a_{k+1}, \ldots, a_n\) activites compatible
                    with one another, and that start after \(a_k\) finishes.
          \end{itemize}
    \item The solutions to the two subproblems must be optimal.
          \begin{itemize}
              \item Prove using the cut-and-paste approach.
          \end{itemize}
\end{itemize}

\subsection*{Optimal Substructure}
\begin{equation*}
    S_{ij} = \text{ subset of activites in } S \text{ that start after } a_i \text{ finishes and finish before } a_j \text{ starts}
\end{equation*}

\begin{align*}
    \text{Let } A_{ik} & = A_{ij} \cap S_{ik} \text{ and } A_{kj} = A_{ij} \cap S_{kj} \\
    A_{ij}             & = A_{ik} \cup \{a_k\} \cup A_{kj}                             \\
    \vert A_{ij} \vert & = \vert A_{ik} \vert + \vert A_{kj} \vert + 1
\end{align*}

\begin{equation*}
    c[i, j] = \text{ size of max-size subset of mutually compatible activites in } S_{ij}
\end{equation*}

\begin{equation*}
    \parbox{5em}{Recursive\\Solution} c[i, j] = \begin{cases}
        0                                                        & \text{if } S_{ij} = \emptyset    \\
        % \max\{c[i, k] + c[k, j] + 1\}                            & \text{if } S_{ij} \neq \emptyset \\
        \underset{a_k \in S_{ij}}{\max}\{c[i, k] + c[k, j] + 1\} & \text{if } S_{ij} \neq \emptyset
    \end{cases}
\end{equation*}

After you choose \(a_k\), the subproblem focuses on the set of
\begin{equation*}
    S_k = \{ a_i \in S : s_i \geq f_k\}
\end{equation*}

\textbf{Theorem}
If \(S_k\) is nonempty and \(a_m\) has the earliest finish time in \(S_k\), then
\(a_m\) is included in some optimal solution.

\textbf{Proof} Let \(A_k\) be an optimal solution to \(S_k\), and let \(a_j\)
have the earliest finish time of any activity in \(A_k\). If \(a_j = a_m\),
done. Otherwise, let \(A_k' = A_k - \{ a_j \} \cup \{ a_m \}\) be \(A_k\) but
with \(a_m\) substituted for \(a_j\).

\begin{verbatim}
Recursive-Activity-Selector(s, f, k, n)
    m = k + 1
    while m <= n and s[m] < f[k]
        m = m + 1
    if m <= n
        return {m} U Recursive-Activity-Selector(s, f, m, n)
    else
        return {}
\end{verbatim}

\subsection*{Typical Steps of Greedy Algorithm}
\begin{itemize}
    \item Cast the optimization problem as one in which we make a (greedy)
          choice and are left with one subproblem to solve.
    \item Prove that there's always an optimal solution to the original problem
          that makes the greedy choice, so that the greedy choice is always
          safe.
    \item Show that greedy choice and optimal solution to the subproblem
          \(\Rightarrow\) optimal solution to the problem.
    \item Make the greedy choice and solve \emph{top-down}.
\end{itemize}
\end{document}